---
title: "ToothGrowth Data Analysis"
author: "Liang Dong"
date: "25 October 2014"
output:
  pdf_document:
    toc: yes
  html_document:
    highlight: tango
    highlights: tango
    keep_md: yes
    theme: readable
    toc: yes
---

# ToothGrowth Data Analysis

## Basic Data Exploratory

First load the library into the R.

```{r setup}
library(ggplot2)
library(datasets)
```

Then use ggplot to plot the ToothGrowth data on different supp.

```{r plot}
ggplot(data=ToothGrowth, aes(x=dose, y=len)) + facet_wrap(~supp) + geom_point() +
    xlab("Dose") + ylab("Tooth Length") + 
    ggtitle("The Effect of Vitamin C on Tooth Growth in Guinea Pigs")
```

## Basic Summary of the Data

The OJ and VC are separated and summarized by different dose level.

```{r summanry}
VC <- subset(ToothGrowth, supp == 'VC')
OJ <- subset(ToothGrowth, supp == 'OJ')

VC$supp <- factor(VC$supp)
OJ$supp <- factor(OJ$supp)

VC05 <- subset(VC, dose==0.5)
VC10 <- subset(VC, dose==1.0)
VC20 <- subset(VC, dose==2.0)
OJ05 <- subset(OJ, dose==0.5)
OJ10 <- subset(OJ, dose==1.0)
OJ20 <- subset(OJ, dose==2.0)

summary(VC05)
summary(OJ05)

summary(VC10)
summary(OJ10)

summary(VC20)
summary(OJ20)
```

## Hypotheis test

From the exploratory analysis and basic summary from first and second section, the mean length of 0.5 and 1.0 dose level from OJ is larger than VC. But on dose level 2.0, the mean length is equal, so it not obvious on which one is higher length.

Base on this initial analysis, suppose the H0 and Ha on the same dose level to be:

* H0: The length of using supplement VC is greater than OJ
* Ha: The alternative hypotheis is length of using supplement VC is no more than method OJ

The same dose level are compared so only three group are compared. Assuming the significant level is 0.05. 

### On dose level 0.5

```{r 0.5}
t.test(OJ05$len, VC05$len, var.equal = FALSE,  alternative = "greater", conf.level = 0.95)
```

The p-value is less than 0.05, so we **reject the null hypothesis**, which means the result is confirming that VC has NO greater impact than OJ on dose level 0.5.

### On dose level 1.0

```{r 1.0}
t.test(OJ10$len, VC10$len, var.equal = FALSE,  alternative = "greater", conf.level = 0.95)
```

The p-value is less than 0.05 as well, so we **reject the null hypothesis** too, which also means the VC has NO greater impact than OJ on dose level 1.0.

### On dose level 2.0

```{r 2.0}
t.test(OJ20$len, VC20$len, var.equal = FALSE,  alternative = "greater", conf.level = 0.95)
```

This time p-value is significant larger than 0.05, so **null hypothesis is NOT rejected**, which confirming the VC has greater impact than OJ on dose level 2.0.

# Conclusion

The conclusion are:

* On dose level 0.5 and 1.0, supplement OJ is better than VC.
* On dose level 2.0, supplement OJ is not better than VC.
